---
title: "Regression Models Course Project"
author: "Vamshi Krishna"
date: "7/12/2020"
output: pdf_document
---

## Executive Summary
  Motor Trend, a magazine about the automobile industry. Below is the data analysis which can answer a question with suitable data.
  1. "Is an automatic or manual transmission better for MPG"
  2. "Quantify the MPG difference between automatic and manual transmissions"

  The below analysis tells us the answers are:
  1. The automatic or manual transmission when compared to MPG for the dataset are not statistically significant.
  2. The MPG difference is 1.8 for automatic vs manual transmissions.
  
## Loading and preprocessing data

Lets load the data and print the head of the datarame and get an idea of columns and data within the dataframe.

```{r load}
data(mtcars)
head(mtcars)
```

We can see that the rownames are the model names and other variables/features are in columns

Lets see the variables types and size of the data
```{r}
print(str(mtcars))
print(dim(mtcars))
```

It is good that all the variables are in number type with 32 rows and 11 columns

## Data cleaning and EDA

By converting some of the variables to factors from numbers, it will be helpful for analysis.
```{r}
columns = c('cyl','vs','am','gear','carb')
mtcars[,columns] <- lapply(mtcars[,columns] , factor)
str(mtcars)
```

Lets see the relation of linear regression between automatic/manual transmissions vs MPG.
the initial assumptions of the model are appropriate mean value and error should be distributed in normal and independent.

```{r}
fit1 = lm(mpg~am,mtcars)
summary(fit1)
#figure1 from appendix
```

lets now look at confidence interval using the formula we know
```{r}
beta_hat = 0.7245
standard_error = 1.764
t_star = qt(1-0.05/2,df = length(mtcars$mpg)-2)
c(beta_hat-t_star*standard_error, beta_hat+t_star*standard_error)
```

From both the plots in figure1, the results of our coefficient summary, small p-value and presence of 0 in CI, we reject the null hypothesis that transmissions affects MPG.

## Multivariate Analysis

Lets now include new variables to increase the standard errors of coefficient estimates of other coorelated regressors.

Lets now create a fit for all the other variables with MPG

```{r}
fit2 = lm(mpg~. , mtcars)
```

Now, we determine which variables are important fot the correlation using stepAIC function.

```{r}
library('MASS')
step = stepAIC(fit2, direction="both", trace=FALSE)
summary(step)
```

The significant variables from the summary in relation to the MPG are cylinders(cyl),horsepower(hp),weight(wt).

## Comparing the models:

To conclude that the secoind model using step is better than the first model, we use ANOVA to compare the models.

```{r}
anova(fit1,step)
```

The difference in both models is the significant. Lets plot the best fit model to view it clearly.(see figure2 from appendix)

## Finding significance of transmission type vs mpg
Lets summarize the coeffients of the best model:

```{r}
coefficients(summary(step))
```

Looking at the p-values in summary data, we can see that the p-value from transmissions vs MPG is not significant. We can prove this using confidence interval as we did above.

```{r}
beta_hat1 = 1.8092
standard_error1 = 1.3963
t_star1 = qt(1-0.05/2, df = length(mtcars$mpg)-2)
c(beta_hat1-t_star1*standard_error1,beta_hat1+t_star1*standard_error1)
```

Since the CI includes 0 and the p-value is .0.5, the difference between automatic transmission and a manual trnasmission does not affect the MPG significantly. However auto-trnasmission is 1.8 greater than the manual.

## Conclusion
By reviewing the above models, the best fit from figure2 shows that the normal Q-Q graph is normally distributed and the scale-location graph has a steady variance which is good when compared to figure1. and finally we can conclude that am does not have a significant impact on mpg.

## Appendix

### Figure 1:

```{r}
par(mfrow=c(2,2))
plot(fit1)
abline(fit1)
```

### Figure 2:
```{r}
par(mfrow=c(2,2))
plot(step)
abline(step)
```

